---
title: "OmicsLake v2.0 Quick Start Guide"
author: "OmicsLake Development Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OmicsLake v2.0 Quick Start Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

OmicsLake v2.0 is a modern, easy-to-use omics data management package with data lineage at its core. It integrates into existing analysis workflows with minimal changes and automatically tracks data dependencies.

## Key Features

- **Simple API**: Basic operations with just `put()`, `get()`, `snap()`, `tree()`
- **Automatic Lineage Tracking**: Dependencies detected through dplyr pipes
- **No SQL Required**: Formula syntax and QueryBuilder for R-native queries
- **Existing Code Support**: observe/wrap modes for non-invasive tracking

# Installation

```{r install}
# Install from GitHub
remotes::install_github("matsui-lab/OmicsLake")
```

# Basic Usage

## Initializing a Lake

```{r init}
library(OmicsLake)

# Create/open a project
lake <- Lake$new("my_analysis")

# Or use global shortcuts
use_lake("my_analysis")
```

## Storing and Reading Data

```{r basic_io}
# Store a data frame
counts <- data.frame(
  gene_id = paste0("GENE", 1:100),
  sample_A = rpois(100, 50),
  sample_B = rpois(100, 60)
)
lake$put("counts", counts)

# R objects can also be stored
params <- list(
  method = "TMM",
  log_transform = TRUE,
  threshold = 0.05
)
lake$put("analysis_params", params)

# Read data
data <- lake$get("counts")
my_params <- lake$get("analysis_params")
```

## Filtering with Formula Syntax

```{r formula_filter}
# No SQL needed! Filter with formula syntax
high_expr <- lake$get("counts", where = ~ sample_A > 50)

# Custom operators are also available
mito_genes <- lake$get("counts", where = ~ gene_id %like% "MT-%")

# Compound conditions
filtered <- lake$get("counts",
  where = ~ sample_A > 30 & sample_B %between% c(40, 80)
)

# Select columns simultaneously
subset <- lake$get("counts",
  where = ~ sample_A > 50,
  select = c("gene_id", "sample_A")
)
```

## dplyr Integration (Automatic Lineage Tracking)

```{r dplyr_integration}
# Get a lazy reference with lake$ref()
# Dependencies in dplyr pipes are automatically tracked!
library(dplyr)

lake$ref("counts") |>
  filter(sample_A > 30) |>
  mutate(
    mean_expr = (sample_A + sample_B) / 2,
    log2_ratio = log2(sample_B / sample_A)
  ) |>
  arrange(desc(mean_expr)) |>
  save_as("processed_counts", lake)

# Check lineage
lake$tree("processed_counts")
# counts -> processed_counts
```

## Joining Multiple Tables

```{r join_example}
# Add metadata
metadata <- data.frame(
  gene_id = paste0("GENE", 1:100),
  gene_name = paste0("Gene_", LETTERS[1:4])[rep(1:4, 25)],
  biotype = sample(c("protein_coding", "lncRNA"), 100, replace = TRUE)
)
lake$put("gene_metadata", metadata)

# JOIN with dplyr
lake$ref("counts") |>
  left_join(lake$ref("gene_metadata"), by = "gene_id") |>
  filter(biotype == "protein_coding") |>
  group_by(gene_name) |>
  summarize(total_A = sum(sample_A), total_B = sum(sample_B)) |>
  save_as("gene_summary", lake)

# Dependencies are automatically tracked
lake$tree("gene_summary")
# counts -----> gene_summary
# gene_metadata ↗
```

# QueryBuilder

Build complex queries without writing SQL.

```{r querybuilder}
# Build queries with method chaining
result <- lake$from("counts")$
  join("gene_metadata", on = "gene_id")$
  where(biotype == "protein_coding")$
  where(sample_A > 40)$
  select(gene_id, gene_name, sample_A, sample_B)$
  order_by(desc(sample_A))$
  top(20, by = sample_A)$
  run()

# Save results to Lake
lake$from("counts")$
  where(sample_A > 50)$
  as("high_expression_genes")
```

# Version Control

## Snapshots and Tags

```{r versioning}
# Snapshot current state
lake$snap("v1.0_raw_data")

# Update data
normalized <- lake$get("counts")
normalized$sample_A <- log2(normalized$sample_A + 1)
normalized$sample_B <- log2(normalized$sample_B + 1)
lake$put("counts", normalized)

lake$snap("v1.1_normalized")

# Tag individual data
lake$tag("counts", "before_normalization")

# Get past version
original <- lake$get("counts", ref = "@tag(before_normalization)")

# Restore to snapshot
lake$restore("v1.0_raw_data")
```

## Viewing History

```{r history}
# Project history
lake$log()

# History of specific data
lake$log("counts")

# List snapshots
lake$snaps()
```

# Lineage (Data Provenance)

## Checking Dependencies

```{r lineage}
# Upstream dependencies (what was this data made from?)
lake$deps("gene_summary", direction = "up")

# Downstream dependencies (what uses this data?)
lake$deps("counts", direction = "down")

# Complete lineage tree
lake$tree("gene_summary", direction = "up", depth = 10)

# Impact analysis (what would be affected if this data changes?)
lake$impact("counts")
```

## Lineage Visualization

```{r plot_lineage}
# Visualize as graph (requires igraph package)
lake$plot("gene_summary", direction = "both")
```

# Lightweight Mode (Integrating with Existing Code)

## observe: Track without Code Changes

```{r observe_mode}
# Run existing scripts in observation mode
result <- observe({
  data <- read.csv("input_data.csv")
  processed <- data[data$value > 0, ]
  write.csv(processed, "output_data.csv")
})

# Files read and written are recorded
print(result$reads)
print(result$writes)
print(result$lineage)
```

## wrap: Wrap Functions for Tracking

```{r wrap_mode}
# Wrap an existing function
normalize_data <- function(x) {
  x$normalized <- scale(x$value)
  x
}

tracked_normalize <- wrap_fn(normalize_data, lake, "normalized_result")

# Use the wrapped function - lineage is automatically recorded
result <- tracked_normalize(my_data)
```

## Defining Pipelines

```{r pipeline}
pipeline <- create_pipeline(lake, "preprocessing")

pipeline$
  step("load", function() read.csv("data.csv"))$
  step("clean", function(data) na.omit(data))$
  step("normalize", function(data) { data$value <- scale(data$value); data })$
  step("filter", function(data) data[data$quality > 0.8, ])

result <- pipeline$run()

# Each step is recorded in the Lake
lake$tree("preprocessing.filter")
```

# Custom Operators

```{r operators}
# %like% - SQL LIKE pattern matching
genes[genes %like% "MT-%"]          # Genes starting with MT-
genes[genes %like% "%kinase%"]      # Contains "kinase"

# %ilike% - Case-insensitive
names[names %ilike% "john%"]

# %between% - Range filter
values[values %between% c(10, 100)]

# %regex% - Regular expression matching
ids[ids %regex% "^ENSG\\d{11}$"]

# %!in% - NOT IN
letters[letters %!in% c("a", "e", "i", "o", "u")]

# is_null / is_not_null
data[is_not_null(data$value), ]
```

# Bracket Notation

```{r bracket_notation}
# Simple read/write
lake["counts"]                      # Read all data
lake["new_data"] <- df              # Write

# Read with filter
lake["counts", sample_A > 50]

# Filter + column selection
lake["counts", sample_A > 50, .(gene_id, sample_A)]
```

# Import/Export

```{r import_export}
# Parquet export
lake$export("counts", "counts.parquet")

# CSV export
lake$export("counts", "counts.csv")

# Import
lake$import("external_data.parquet", "imported_data")
lake$import("annotations.csv", "gene_annotations")
```

# Bioconductor Integration

```{r bioconductor}
library(SummarizedExperiment)

# Store SummarizedExperiment directly
# assays, colData, rowData, metadata are all preserved
lake$put("rna_experiment", se_object)

# Restore completely
se_restored <- lake$get("rna_experiment")
```

# Global Shortcuts

```{r shortcuts}
# Set default Lake
use_lake("my_project")

# Operate without lake$ prefix
put("data", df)
data <- fetch("data")      # fetch() instead of get()
snap("checkpoint1")
tree("data")
tables()
history()
```

# Migration from Legacy API

The v1.0 `ol_*` functions continue to work, but migration to the new API is recommended:

```{r migration}
# Show migration guide
show_migration_guide()
```

| Legacy API | New API |
|------------|---------|
| `ol_init("proj")` | `Lake$new("proj")` |
| `ol_write("t", df)` | `lake$put("t", df)` |
| `ol_read("t")` | `lake$get("t")` |
| `ol_label("v1")` | `lake$snap("v1")` |
| `ol_tag("t", "v1")` | `lake$tag("t", "v1")` |
| `ol_checkout("v1")` | `lake$restore("v1")` |
| `ol_show_lineage("t")` | `lake$tree("t")` |
| `ol_query("SQL")` | `lake$sql("SQL")` |

# Summary

Key features of OmicsLake v2.0:

1. **Lake R6 Class** - Simple and intuitive API
2. **Automatic Lineage Tracking** - Dependencies detected through dplyr pipes
3. **QueryBuilder** - Query construction without SQL
4. **Formula Syntax** - R-like filter notation
5. **Lightweight Mode** - Non-invasive integration with existing code
6. **Bioconductor Integration** - Full SE/MAE support

See individual function help for details:

```{r help}
?Lake
?QueryBuilder
?observe
?wrap_fn
```
